### 2018-04-03 Datelife_paper1.Rmd
missing references

DateLife also stores information on input taxon presence/absence across subset chronograms. Users can choose to add ages of missing taxa to subset chronograms in different ways, depending on the amount of knowledge or how much they want to be involved in the steps of the addition process.
If users have no access to biological information (i.e., a character, DNA or protein matrix), missing taxa can be added to any chronogram simply at random, or by following taxonomic or phylogenetic knowledge from expert sources. There are a wide number of open reference taxonomies available, such as the Catalogue of Life [@Roskob2017] or the NCBI taxonomy database [@Federhen2012]. Expert phylogenies (with or without branch lengths) to be used as topological constraint (backbone) can also be obtained from a number of public repositories, such as OToL [@Hinchliff2015], TreeBASE [@Piel2002] and Dryad (<https://www.datadryad.org//>).
At the moment, DateLife only uses OToL's synthetic tree and reference taxonomy as expert knowledge to automatically add missing taxa to chronograms. Alternatively, users can input a reference taxonomy or topological constraint of their choosing or making. <!-- at some point we will implement the possibility for users to choose different reference taxonomies automatically-->
If OToL's synthetic tree is not satisfactorily resolved for the taxa of interest, DateLife can construct a sequence data matrix from DNA markers available from the Barcode of Life Database (BOLD; @Ratnasingham2007), to attempt to further resolve polytomies. It will follow OToL's synthetic tree as backbone. <!--This molecular tree can then be dated with chronosMPL from the ape package, but is it really needed?? To estimate node ages chronosMPL uses mean path length method from Britton et al. 2002-->
<!--It can rapidly construct trees with branch lengths from
a set of lineages with sequence data from the Barcode Of Life Database (BOLD, <http://www.boldsystems.org/>), using the synthetic OToL (<https://tree.opentreeoflife.org>) as backbone, and to obtain divergence dates.
It allows direct comparison of dates obtained with different markers
available in BOLD (in plants and fungi in particular).-->
To use information from a topological constraint, DateLife calls the congruification method described in [@Eastman2013] to find shared nodes between trees (congruent nodes). It then fixes their ages, and add ages to remaining nodes with a dating method that can be specified by the user. If users have access to biological data, thay can input a tree with branch lengths proportional to relative substitution rates as topological constraint. In this case, age data from congruent nodes will be used as calibration points.
Age data from several chronograms can be combined and congruified to be used as calibration points in a single analysis.

>>>>

DateLife also stores information on input taxon presence/absence across subset chronograms. Users can choose to add ages of missing taxa to subset chronograms in different ways, depending on the amount of knowledge or how much they want to be involved in the steps of the addition process.
If users have no access to biological information (i.e., a character, DNA or protein matrix), missing taxa can be added to any chronogram simply at random, or by following taxonomic or phylogenetic knowledge from expert sources. There are a wide number of open reference taxonomies available, such as the Catalogue of Life [@Roskov2017] or the NCBI taxonomy database [@Federhen2012]. Expert phylogenies (with or without branch lengths) to be used as topological constraint (backbone) can also be obtained from a number of public repositories, such as OToL [@Hinchliff2015], TreeBASE [@Piel2002] and Dryad (<https://www.datadryad.org//>).
At the moment, DateLife only uses OToL's synthetic tree and reference taxonomy as expert knowledge to automatically add missing taxa to chronograms. Alternatively, users can input a reference taxonomy or topological constraint of their choosing or making. <!-- at some point we will implement the possibility for users to choose different reference taxonomies automatically-->
If OToL's synthetic tree is not satisfactorily resolved for the taxa of interest, DateLife can construct a sequence data matrix from DNA markers available from the Barcode of Life Database (BOLD; @Ratnasingham2007), to attempt to further resolve polytomies. It will follow OToL's synthetic tree as backbone. <!--This molecular tree can then be dated with chronosMPL from the ape package, but is it really needed?? To estimate node ages chronosMPL uses mean path length method from Britton et al. 2002-->
<!--It can rapidly construct trees with branch lengths from
a set of lineages with sequence data from the Barcode Of Life Database (BOLD, <http://www.boldsystems.org/>), using the synthetic OToL (<https://tree.opentreeoflife.org>) as backbone, and to obtain divergence dates.
It allows direct comparison of dates obtained with different markers
available in BOLD (in plants and fungi in particular).-->
To use information from a topological constraint, DateLife calls the congruification method described in [@Eastman2013] to find shared nodes between trees (congruent nodes). It then fixes their ages, and add ages to remaining nodes with a dating method that can be specified by the user. If users have access to biological data, thay can input a tree with branch lengths proportional to relative substitution rates as topological constraint. In this case, age data from congruent nodes will be used as calibration points.
Age data from several chronograms can be combined and congruified to be used as calibration points in a single analysis.
